Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Edit responsiveness results

Let's focus on the left figure which illustrates a work stealing
example: when the non-cooperative kernel is given only one workgroup, it
execution is so long that it cannot permit to terminate on time for a
screen refresh. The gather time is very good though, since the scheduler
only needs to collect one workgroup. As more workgroups are allocated to
the non-cooperative kernel, it can benefits from more computational
resources and thus lower its execution time: here the non-cooperative
kernel becomes fast enough with a quarter (resp. half) of available
workgroups for light (resp. heavy) graphics workload. Inversely, the
gather time augments since the scheduler must collect more and more
workgroups from the cooperative kernel.

On the middle and right figures, which depict graph algorithms, we find
again the same trend: as the number of workgroups allocated to the
non-cooperative task grows, the non-cooperative kernel execution time
lowers and the gather time raises. When the regular resizing barrier is
used, gather time can become quite long and exceed the execution time.
However, the scheduling responsiveness can benefits from the query
barrier: when this barrier is used, the gather time becomes almost
stable w.r.t. the number of workgroups that needs to be collected, since
the scheduler collects all needed workgroups as soon as they hit a
resizing barrier.

\HEComment{missing results for all but one in bfs, discussed in expe
setup, do we need to say it here ?}

@@

There are three main avenues for future work at present: seeking
additional classes of irregular algorithms to which the model might
(be extended to) apply (to), investigating native driver support for
cooperative kernels via open source graphics drivers, and integrating
cooperative kernels into template- and compiler-based programming
models for graph algorithms on
GPUs~\cite{DBLP:conf/ppopp/WangDPWRO16,DBLP:conf/oopsla/PaiP16}.  As more vendors
adopt the advanced features of OpenCL that our approach requires, we
will be in a position to assess the programming model across a wider
range of devices.

>>>>

Let's focus on the left figure which illustrates a work stealing
example: when the non-cooperative kernel is given only one workgroup, it
execution is so long that it cannot permit to terminate on time for a
screen refresh. The gather time is very good though, since the scheduler
only needs to collect one workgroup. As more workgroups are allocated to
the non-cooperative kernel, it can benefits from more computational
resources and thus lower its execution time: here the non-cooperative
kernel becomes fast enough with a quarter (resp. half) of available
workgroups for light (resp. heavy) graphics workload. Inversely, the
gather time augments since the scheduler must collect more and more
workgroups.

The middle and right figures, which depict graph algorithms, expose the
same trend: as the number of non-cooperative workgroups grows, the
execution time lowers and the gather time raises.\footnote{The right
figure has only 3 points per line: recall that for bfs there are no runs
with ``all but one'' non-cooperative workgroups}
%
The gather time is higher on the right figure: the rmat22 input graph is
rather wide than deep, so the graph algorithm reaches resizing barriers
less often than for the ecology input of the middle figure for
instance. The scheduler has thus fewer opportunities to collect
workgroups and gather time increases. Nonetheless, scheduling
responsiveness can benefit from the query barrier: when used, this
barrier let the scheduler collects all needed workgroups as soon as they
hit a resizing barrier. As we can see on both the middle and right
figure, the gather time of the query barrier is almost stable w.r.t. the
number of workgroups that needs to be collected.

@@

There are three main avenues for future work at present: seeking
additional classes of irregular algorithms to which the model might
(be extended to) apply (to), investigating native driver support for
cooperative kernels via open source graphics drivers, and integrating
cooperative kernels into template- and compiler-based programming
models for graph algorithms on
GPUs~\cite{SreesCompetition,DBLP:conf/oopsla/PaiP16}.  As more vendors
adopt the advanced features of OpenCL that our approach requires, we
will be in a position to assess the programming model across a wider
range of devices.
